Verifying Hierarchical Cache Coherence Protocols: a theorem-prover based modular, reusable proof

### 2014-02-03 paper/Intuition.tex
some modifications of background

\begin{inv}
\textit{stateChange}:
\begin{enumerate}
\item A child sends a response to its parent iff it downgrades its state.
\item A parent sends a response to a child iff it upgrades the directory of
that child.
\item The state and the directory can change only on sending and receiving a
response.
\end{enumerate}
\label{stateChange}
\end{inv}

@@

@@

Using Invariants \ref{stateChange} to \ref{noTransitEqual}, it is easy to prove
Invariant \textit{conservative}. If no response is in flight for an address, then
the parent's directory state of the cache for the address matches that of the
cache's state for that address. If there is a response from parent to child in
flight, then the corresponding directory at the parent will be no less than the
directory immediately after the response was sent and the state of the child
will be no greater than the directory immediately before the response was sent
(by Invariant \textit{inTransitCondition}). By Invariant \textit{stateChange}, the
parent upgrades the directory when sending the response, thus conforming to
Invariant \textit{conservative}. Similarly, it can be shown that if there is a
response from child to parent, Invariant \textit{conservative} will hold.

\begin{inv}
\textit{respFIFO}: Responses are received in the order in which they are sent,
for a source-destination pair.
\label{respFIFO}
\end{inv}

>>>>

\begin{inv}
\textit{stateChange}:
\begin{enumerate}
\item A child sends a response to its parent iff it downgrades its state.
\item A parent sends a response to a child iff it upgrades the directory of
that child.
\item The state and the directory can change only on sending and receiving a
response; the receiving entity changes its value to that of the sending entity
immediately after sending the response message.
\item 
\end{enumerate}
\label{stateChange}
\end{inv}

\begin{inv}
\textit{respFIFO}: Responses are received in the order in which they are sent,
for a source-destination pair, for the same address.
\label{respFIFO}
\end{inv}

Invariant \textit{conservative} at any time $t$ for child cache $c$ and parent
$p$ for an address $a$ can be proved as follows. Consider the case when
$c.\state[a]$ never changed before $t$, then $c.\state[a] = \In$ because of the
initialization conditions described in Section \ref{System}. No response would
have been sent by $c$ to its parent for address $a$ before $t$, otherwise
$c.\state[a]$ would have changed. Thus the parent would not have received any
response from $c$, $p.\dstate[c][a]$ could have changed only by sending a
response to $c$ which can only upgrade it from the initial value of $\In$. Thus
at $t$, $p.\dstate[c][a] \ge c.state[a]$.

Now consider the case when $c.\state[a]$ last changed at $t'$ such that $t' <
t$ because of receiving a response from $p$, sent at $t''$. $c.\state[a]$ at
$t'+1 = p.\dstate[c][a]$ at $t''+1$ because of Invariant \textit{stateChange}.
$c$ neiter sends nor receives a response at any $t_i$ such that $t' < t_i < t$.
If $p$ received a response at some $t_m$ such that $t'' < t_m < t$,
then it would either violate Invariant \textit{noCross}, or the fact that no
message was sent between $t' < t_i < t$ from $c$. Thus, $p.\dstate[c][a]$ at $t
\ge p.\dstate[c][a]$ at $t''+1$ because any message sent can only upgrade
$p.\dstate[a]$. This ensures that Invariant \textit{conservative} is obeyed. We
can similarly prove this invariant holds when $c.\state[a]$ last changes before
$t$ by sending a response.

@@

Invariant \textit{respFIFO} can be enforced by the interconnect network which
ensures a FIFO ordering for messages sent between a pair of nodes.

%Invariant \textit{respFIFO} is known as point-to-point FIFO ordering in the
%Network-on-Chip NoC community, and is enforced by common routing algorithms
%like Dimension-ordered routing \cite{dimension}. Only the parent-to-child
%channels need to be FIFO ordered, the FIFO ordering in child-to-parent channels
%can be enforced using the protocol itself.

---------------------------------------------

@@
